 ---
title: "Curator Desk"
format:
  html:
    toc: false
---

# BioAnalyzer curator desk

Use this page to review BioAnalyzer’s predicted statuses for candidate papers and record curator ground truth for each PMID.

What you can do here:

* **Find papers quickly:** Search, sort, and filter the table (Priority score helps surface high‑value candidates).
* **Record ground truth:** Enter curator TRUE labels and optional “was the prediction correct?” flags per field.
* **Share feedback:** Submit feedback via a GitHub issue (CSV is pre-filled in the issue body).

::: {.callout-note collapse="false"}
## Data Source & Setup

The table is built from data stored in the **`data/`** folder. To analyze your own papers: 

1.  Place your CSV (or Parquet) file in `data/` (e.g., `data/sample.csv`).
2.  Run `quarto render`. 

The system reads `data/sample.csv` by default. To use a different source, set the `CURATOR_DATA_PATH` environment variable to your specific path (e.g., `data/analyzed_papers.csv`) or use `CURATOR_DATA_URL` for remote files.

**Expected columns**

* **Required:** `PMID`
* **Recommended:** `Title`, `Journal`, `Year`, `Summary`
* **Status values:** `ABSENT` | `PARTIALLY_PRESENT` | `PRESENT`
:::

---

## Candidate Curatable Articles {.tabset #candidate-papers}

::: {.panel-tabset}

### Interactive Table

**Tip:** Sort by **Priority** (descending) to review the most promising candidates first. Use the search box and column filters to narrow your focus.

```{r echo = FALSE}
#| label: setup
#| include: false

# ---- Load dependencies ----
source("R/config.R")
source("R/data.R")
source("R/feedback.R")

# ---- Resolve data source ----

resolve_data_path <- function() {
  path <- Sys.getenv("CURATOR_DATA_URL", "")
  if (!nzchar(trimws(path))) path <- Sys.getenv("CURATOR_DATA_PATH", "")
  if (!nzchar(trimws(path))) path <- "data/sample.csv"
  path
}

data_path <- resolve_data_path()

# ---- Load and normalize dataset ----

raw_df <- load_data(data_path)
df <- normalize_dataset(raw_df)

# ---- Validate expected status columns ----

missing_status <- setdiff(STATUS_COLUMNS, names(df))

if (length(missing_status) > 0) {
  warning(
    sprintf(
      "Missing expected status columns: %s. Priority and filtering may be partial.",
      paste(missing_status, collapse = ", ")
    )
  )
}
```
:::


```{r}
#| label: table-display
#| echo: false
#| warning: false
#| message: false
#| column: screen

library(dplyr)

if (nrow(df) == 0) {
  
  cat(
    "No data loaded. Set `CURATOR_DATA_URL` or `CURATOR_DATA_PATH`, ",
    "or add `data/sample.csv` and rebuild."
  )
  
} else {
  
  # ---- Inform about missing status columns ----
  
  if (length(missing_status) > 0) {
    cat(
      "**Note:** Missing status columns: ",
      paste(missing_status, collapse = ", "),
      ". Priority and filtering may be partial.\n\n"
    )
  }
  
  # ---- Select display columns ----
  
  DISPLAY_COLUMNS <- c(
    "PubMed Link",
    "Title",
    "Journal",
    "Year",
    "Host Species",
    "Host Species Status",
    "Body Site",
    "Body Site Status",
    "Condition",
    "Condition Status",
    "Sequencing Type",
    "Sequencing Type Status",
    "Sample Size",
    "Sample Size Status",
    "has_differential_abundance",
    "differential_abundance_confidence",
    "in_bugsigdb",
    "Priority"
  )
  display_columns <- intersect(DISPLAY_COLUMNS, names(df))
  df_display <- df[, display_columns, drop = FALSE]

  # ---- Configure DataTable options ----
  
 dt_options <- list(
  pageLength = 50,
  lengthMenu = c(50, 100, 300, 500, 1000),
  scrollX = TRUE,
  stateSave = TRUE,
  fixedHeader = TRUE
)
  
  priority_index <- which(names(df_display) == "Priority")
  
  if (length(priority_index) == 1) {
    dt_options$order <- list(list(priority_index - 1, "desc"))
  }
  
  in_bugsigdb_index <- which(names(df_display) == "in_bugsigdb")
  dt_column_defs <- list()
  if (length(in_bugsigdb_index) == 1) {
    dt_column_defs <- append(
      dt_column_defs,
      list(
        list(
          targets = in_bugsigdb_index - 1,
          title = "In BugSigDB",
          render = DT::JS(
            "function(data, type, row) {",
            "  if (type === 'display') {",
            "    return data ? '<span class=\"bugsigdb-yes\">✓ Yes</span>' : '<span class=\"bugsigdb-no\">✗ No</span>';",
            "  }",
            "  return data;",
            "}"
          )
        )
      )
    )
  }

  dt_options$columnDefs <- dt_column_defs

  # ---- Render table ----
  dt_table <- DT::datatable(
    df_display,
    filter = "top",
    extensions = c("FixedHeader"),,
    options = dt_options,
    escape = FALSE,
    rownames = FALSE
  )

  if ("in_bugsigdb" %in% names(df_display)) {
    dt_table <- dt_table %>%
      DT::formatStyle(
        "in_bugsigdb",
        backgroundColor = DT::styleEqual(c(TRUE, FALSE), c("#e8f5e9", "#ffffff"))
      )
  }

  for (status_col in STATUS_COLUMNS) {
    if (status_col %in% names(df_display)) {
      dt_table <- dt_table %>%
        DT::formatStyle(
          status_col,
          backgroundColor = DT::styleEqual(
            c("PRESENT", "PARTIALLY_PRESENT", "ABSENT"),
            c("#c8e6c9", "#fff9c4", "#ffcdd2")
          )
        )
    }
  }

  dt_table
}
```
::: {.panel-tabset}

**Rows in table:** `r nrow(df)`

## Curator feedback

Use the table above to find papers (the **Open** button in the **Published** column opens the paper on PubMed). Then select or enter a PMID below; the page will scroll to the field-by-field validation form where you can enter your review. Feedback uses the same schema as the Streamlit app; because this site is static, **saving is done by downloading a CSV** per row—submit via your usual process (e.g. email or GitHub issue).

1. **Quick select** a PMID from the dropdown, or type a PMID and click **Load row** (the form will scroll into view).
2. For each field, review **BioAnalyzer predicted** and set **Curator TRUE label** and **Was BioAnalyzer correct?**
3. Click **Download feedback CSV** to save one row; submit as needed.

```{r}
#| label: table-json
#| echo: false
#| results: asis
# Export table rows as JSON for the form's "Load row" button (PMID -> pred values)
if (nrow(df) > 0) {
  want <- c("PMID", "Title", STATUS_COLUMNS)
  export_cols <- intersect(want, names(df))
  rows_json <- jsonlite::toJSON(
    df[, export_cols, drop = FALSE],
    auto_unbox = TRUE,
    dataframe = "rows"
  )
  cat(sprintf('<script type="application/json" id="curator-table-data">%s</script>', rows_json))
}
```

```{r}
#| label: curator-config
#| echo: false
#| results: asis
# Defaults for form (same as Python CONFIG)
cat(sprintf('<script type="application/json" id="curator-config">%s</script>',
            jsonlite::toJSON(list(curator_id_default = CONFIG$curator_id_default,
                                 bioanalyzer_version_default = CONFIG$bioanalyzer_version_default,
                                 feedback_issue_url = CONFIG$feedback_issue_url),
                            auto_unbox = TRUE)))
```

::: {#feedback-form .card .shadow-sm}
::: {.card-header .bg-light .py-3}
**Curator feedback form**
:::
::: {.card-body}
::: {.mb-3}
**Curator ID / initials**  
<input type="text" id="curator_id" class="form-control" placeholder="e.g. Ronald Ouma" style="max-width: 320px;">
:::
::: {.mb-3}
**Quick select for feedback** — Select a PMID from the table, or type below.
:::

```{r}
#| label: quick-select-pmid
#| echo: false
#| results: asis
if (nrow(df) > 0 && "PMID" %in% names(df)) {
  opts <- df[, c("PMID", if ("Title" %in% names(df)) "Title" else NULL), drop = FALSE]
  cat('<select id="quick-select-pmid" class="form-select" style="max-width: 480px;"><option value="">— Select a PMID —</option>')
  for (i in seq_len(nrow(opts))) {
    pm <- opts$PMID[i]
    lbl <- if ("Title" %in% names(opts)) paste0(pm, " — ", substr(opts$Title[i], 1, 50), if (nchar(as.character(opts$Title[i])) > 50) "…" else "") else as.character(pm)
    lbl <- gsub("&", "&amp;", gsub("<", "&lt;", gsub(">", "&gt;", lbl)))
    cat(sprintf('<option value="%s">%s</option>', pm, lbl))
  }
  cat('</select>')
}
```

::: {.mb-3}
**PMID**  
<div class="input-group" style="max-width: 360px;">
  <input type="text" id="fb_pmid" class="form-control" placeholder="e.g. 31215600">
  <button type="button" id="btn-load-row" class="btn btn-outline-secondary">Load row</button>
</div>
:::
::: {.mb-3}
**Title (from table)**  
<span id="title-prefill" class="text-muted fst-italic"></span>
:::

::: {.mb-3}
**Overall paper verdict**  
<select id="overall_verdict" class="form-select" style="max-width: 280px;">
  <option value="Curatable">Curatable</option>
  <option value="Not curatable">Not curatable</option>
  <option value="Uncertain">Uncertain</option>
  <option value="Not reviewed">Not reviewed</option>
</select>
:::
::: {.mb-3}
**Comment (optional)**  
<textarea id="comment" class="form-control" rows="3" placeholder="Evidence, edge case, missing field, false positive reason..." style="max-width: 520px;"></textarea>
:::
::: {.mb-3}
**BioAnalyzer version** (fixed for all submissions)  
<input type="text" id="bioanalyzer_version" class="form-control" readonly style="max-width: 200px; background-color: var(--bs-light, #f8f9fa);" title="Version is set at build time.">
:::

### Field-by-field validation (ground truth) {#field-validation-section}

For each field, provide the **curator TRUE label** (ground truth). Optionally mark whether BioAnalyzer's predicted status was correct.

::: {.grid}

```{r}
#| label: feedback-fields
#| echo: false
#| results: asis
for (col in STATUS_COLUMNS) {
  s <- safe_col(col)
  label <- gsub(" Status", "", col)
  cat(sprintf('
<div class="feedback-field">
  <strong>%s</strong>
  <p class="bioanalyzer-pred-line"><span id="pred_%s" class="bioanalyzer-pred text-muted">Select a PMID above and click Load row to see BioAnalyzer&rsquo;s prediction.</span></p>
  <small>Curator TRUE label for %s:</small>
  <select id="true_%s">
    <option value="Not reviewed">Not reviewed</option>
    <option value="ABSENT">ABSENT</option>
    <option value="PARTIALLY_PRESENT">PARTIALLY_PRESENT</option>
    <option value="PRESENT">PRESENT</option>
  </select>
  <small>Was BioAnalyzer correct for %s?</small>
  <select id="col_fb_%s">
    <option value="Not reviewed">Not reviewed</option>
    <option value="Correct">Correct</option>
    <option value="Incorrect">Incorrect</option>
    <option value="Unclear">Unclear</option>
  </select>
</div>
', label, s, label, s, label, s))
}
```

:::

**Submit review** — Opens GitHub to create an issue with your feedback (no file download). A bot will then open a Pull Request with the feedback CSV so the maintainer can view and download it (including your initials and comments).

::: {.mt-2}
<button type="button" id="btn-submit-review" class="btn btn-primary me-2">Submit review</button>
<button type="button" id="btn-download-csv" class="btn btn-outline-secondary">Download CSV only</button>
:::

::: {.mt-3 .pt-3 .border-top .small .text-muted}
**Notes:** Submissions are processed as issues and a tab will be opened for you to just click the create button. 
Reviewers will be looking into you submittd issue to evaluate the results
:::
:::
:::

::: {#thank-you-modal .thank-you-overlay}
::: {.thank-you-card}
**Thank you** for your time and input in helping to complete this review.  
Your feedback has been submitted; an issue and pull request are being created in the background.

Would you like to continue reviewing other papers?

::: {.thank-you-actions .mt-3}
<button type="button" id="btn-continue-review" class="btn btn-primary me-2">Yes</button>
<button type="button" id="btn-go-to-index" class="btn btn-outline-secondary">No</button>
:::
:::
:::

<script src="js/feedback-form.js"></script>

<style>
.thank-you-overlay {
  position: fixed;
  top: 0; left: 0; right: 0; bottom: 0;
  background: rgba(0,0,0,0.5);
  display: flex;
  align-items: center;
  justify-content: center;
  z-index: 9999;
  opacity: 0;
  visibility: hidden;
  transition: opacity 0.25s ease, visibility 0.25s ease;
}
.thank-you-overlay.show {
  opacity: 1;
  visibility: visible;
}
.thank-you-card {
  background: var(--bs-body-bg, #fff);
  padding: 2rem;
  border-radius: 0.5rem;
  box-shadow: 0 0.5rem 2rem rgba(0,0,0,0.2);
  max-width: 420px;
  text-align: center;
}
.thank-you-actions { display: flex; justify-content: center; flex-wrap: wrap; gap: 0.5rem; }

#feedback-form .card-body { padding: 1.25rem 1.5rem; }
.feedback-field {
  margin-bottom: 1.25rem;
  padding: 0.75rem;
  background: var(--bs-light, #f8f9fa);
  border-radius: 0.35rem;
  border: 1px solid rgba(0,0,0,.06);
}
.feedback-field strong { display: block; margin-bottom: 0.35rem; }
.feedback-field .bioanalyzer-pred-line { margin: 0.35rem 0 0.5rem 0; min-height: 1.4em; }
.bioanalyzer-pred { font-size: 0.95em; font-weight: 500; }
.bioanalyzer-pred.pred-PRESENT { color: #198754; }
.bioanalyzer-pred.pred-ABSENT { color: #dc3545; }
.bioanalyzer-pred.pred-PARTIALLY_PRESENT { color: #fd7e14; }
.feedback-field select { margin-right: 0.5rem; margin-top: 0.25rem; max-width: 12em; }
.feedback-field small { color: var(--bs-secondary); }
@media (min-width: 768px) {
  .grid { display: grid; grid-template-columns: 1fr 1fr; gap: 1rem 1.5rem; }
}
</style>
