 ---
title: "Curator Desk"
format:
  html:
    toc: false
---

# BioAnalyzer Curation Hub: Turning Predictions into Progress

Welcome to your central workspace for refining PubMed literature discovery. This dashboard bridges the gap between AI-driven BioAnalyzer predictions and expert human insight, putting a searchable, high-performance table of candidate articles at your fingertips.

As you review and capture ground truth for each PMID, the system works in the background to store statuses and track correctness flags in real-time. This seamless integration of human feedback allows you to:

* **Streamline Validation:** Effortlessly navigate and filter high-priority candidates.
* **Refine Model Accuracy:** Log curator ground truth to build a robust data foundation.
* **Generate High-Level Analytics:** Export data ready for deep-dive error profiling, confusion matrices, and MCC performance metrics to help evaluate BioAnalyzer package accuracy.

::: {.callout-note collapse="false"}
## Data Source & Setup

The table is built from data stored in the **`data/`** folder. To analyze your own papers: 

1.  Place your CSV (or Parquet) file in `data/` (e.g., `data/sample.csv`).
2.  Run `quarto render`. 

The system reads `data/sample.csv` by default. To use a different source, set the `CURATOR_DATA_PATH` environment variable to your specific path (e.g., `data/analyzed_papers.csv`) or use `CURATOR_DATA_URL` for remote files.

**Expected Columns:** * **Required:** `PMID` 
* **Recommended:** `Title`, `Journal`, `Year`, `Summary`
* **Status Values:** `ABSENT` | `PARTIALLY_PRESENT` | `PRESENT`
:::

---

## Candidate Curatable Articles {.tabset #candidate-papers}

::: {.panel-tabset}

### Interactive Table

**Tip:** Sort by **Priority Score** (descending) to review the most promising candidates first. Use the search box and column filters to narrow your focus.

```{r}
#| echo: false
# Your code to render the table (DT or reactable) goes here
```{r}
#| label: setup
#| include: false

# ---- Load dependencies ----
source("R/config.R")
source("R/data.R")
source("R/feedback.R")

# ---- Resolve data source ----

resolve_data_path <- function() {
  path <- Sys.getenv("CURATOR_DATA_URL", "")
  if (!nzchar(trimws(path))) path <- Sys.getenv("CURATOR_DATA_PATH", "")
  if (!nzchar(trimws(path))) path <- "data/sample.csv"
  path
}

data_path <- resolve_data_path()

# ---- Load and normalize dataset ----

raw_df <- load_data(data_path)
df <- normalize_dataset(raw_df)

# ---- Validate expected status columns ----

missing_status <- setdiff(STATUS_COLUMNS, names(df))

if (length(missing_status) > 0) {
  warning(
    sprintf(
      "Missing expected status columns: %s. Priority and filtering may be partial.",
      paste(missing_status, collapse = ", ")
    )
  )
}
```

```{r}
#| label: table-display
#| echo: false
#| warning: false
#| message: false

if (nrow(df) == 0) {
  
  cat(
    "No data loaded. Set `CURATOR_DATA_URL` or `CURATOR_DATA_PATH`, ",
    "or add `data/sample.csv` and rebuild."
  )
  
} else {
  
  # ---- Inform about missing status columns ----
  
  if (length(missing_status) > 0) {
    cat(
      "**Note:** Missing status columns: ",
      paste(missing_status, collapse = ", "),
      ". Priority and filtering may be partial.\n\n"
    )
  }
  
  # ---- Select display columns ----
  
  desired_columns <- c(
    "PMID",
    "PubMed Link",
    "Priority Score",
    "Title",
    "Journal",
    "Year",
    STATUS_COLUMNS,
    "Summary"
  )
  
  display_columns <- intersect(desired_columns, names(df))
  show_df <- df[, display_columns, drop = FALSE]
  
  # ---- Convert PubMed Link to clickable column ----
  
  if ("PubMed Link" %in% names(show_df)) {
    
    show_df$PubMed <- paste0(
      '<a href="',
      show_df[["PubMed Link"]],
      '" target="_blank" rel="noopener">Open</a>'
    )
    
    show_df[["PubMed Link"]] <- NULL
    
    # Place PubMed column immediately after PMID; label as "Published" (Open = link to paper on PubMed)
    if ("PMID" %in% names(show_df)) {
      show_df <- show_df[
        ,
        c("PMID", "PubMed",
          setdiff(names(show_df), c("PMID", "PubMed"))
        ),
        drop = FALSE
      ]
    }
    names(show_df)[names(show_df) == "PubMed"] <- "Published"
  }
  
  # ---- Configure DataTable options ----
  
  dt_options <- list(
    pageLength = 50,
    lengthMenu = c(50, 100, 300, 500, 1000),
    scrollX = TRUE,
    stateSave = TRUE
  )
  
  priority_index <- which(names(show_df) == "Priority Score")
  
  if (length(priority_index) == 1) {
    dt_options$order <- list(list(priority_index - 1, "desc"))
  }
  
  # ---- Render table ----
  
  DT::datatable(
    show_df,
    filter = "top",
    options = dt_options,
    escape = FALSE,
    rownames = FALSE
  )
}
```

**Rows in table:** `r nrow(df)`

## Curator feedback

Use the table above to find papers (the **Open** button in the **Published** column opens the paper on PubMed). Then select or enter a PMID below; the page will scroll to the field-by-field validation form where you can enter your review. Feedback uses the same schema as the Streamlit app; because this site is static, **saving is done by downloading a CSV** per row—submit via your usual process (e.g. email or GitHub issue).

1. **Quick select** a PMID from the dropdown, or type a PMID and click **Load row** (the form will scroll into view).
2. For each field, review **BioAnalyzer predicted** and set **Curator TRUE label** and **Was BioAnalyzer correct?**
3. Click **Download feedback CSV** to save one row; submit as needed.

```{r}
#| label: table-json
#| echo: false
#| results: asis
# Export table rows as JSON for the form's "Load row" button (PMID -> pred values)
if (nrow(df) > 0) {
  want <- c("PMID", "Title", STATUS_COLUMNS)
  export_cols <- intersect(want, names(df))
  rows_json <- jsonlite::toJSON(
    df[, export_cols, drop = FALSE],
    auto_unbox = TRUE,
    dataframe = "rows"
  )
  cat(sprintf('<script type="application/json" id="curator-table-data">%s</script>', rows_json))
}
```

```{r}
#| label: curator-config
#| echo: false
#| results: asis
# Defaults for form (same as Python CONFIG)
cat(sprintf('<script type="application/json" id="curator-config">%s</script>',
            jsonlite::toJSON(list(curator_id_default = CONFIG$curator_id_default,
                                 bioanalyzer_version_default = CONFIG$bioanalyzer_version_default,
                                 feedback_issue_url = CONFIG$feedback_issue_url),
                            auto_unbox = TRUE)))
```

::: {#feedback-form .card .shadow-sm}
::: {.card-header .bg-light .py-3}
**Curator feedback form**
:::
::: {.card-body}
::: {.mb-3}
**Curator ID / initials**  
<input type="text" id="curator_id" class="form-control" placeholder="e.g. Ronald Ouma" style="max-width: 320px;">
:::
::: {.mb-3}
**Quick select for feedback** — Select a PMID from the table, or type below.
:::

```{r}
#| label: quick-select-pmid
#| echo: false
#| results: asis
if (nrow(df) > 0 && "PMID" %in% names(df)) {
  opts <- df[, c("PMID", if ("Title" %in% names(df)) "Title" else NULL), drop = FALSE]
  cat('<select id="quick-select-pmid" class="form-select" style="max-width: 480px;"><option value="">— Select a PMID —</option>')
  for (i in seq_len(nrow(opts))) {
    pm <- opts$PMID[i]
    lbl <- if ("Title" %in% names(opts)) paste0(pm, " — ", substr(opts$Title[i], 1, 50), if (nchar(as.character(opts$Title[i])) > 50) "…" else "") else as.character(pm)
    lbl <- gsub("&", "&amp;", gsub("<", "&lt;", gsub(">", "&gt;", lbl)))
    cat(sprintf('<option value="%s">%s</option>', pm, lbl))
  }
  cat('</select>')
}
```

::: {.mb-3}
**PMID**  
<div class="input-group" style="max-width: 360px;">
  <input type="text" id="fb_pmid" class="form-control" placeholder="e.g. 31215600">
  <button type="button" id="btn-load-row" class="btn btn-outline-secondary">Load row</button>
</div>
:::
::: {.mb-3}
**Title (from table)**  
<span id="title-prefill" class="text-muted fst-italic"></span>
:::

::: {.mb-3}
**Overall paper verdict**  
<select id="overall_verdict" class="form-select" style="max-width: 280px;">
  <option value="Curatable">Curatable</option>
  <option value="Not curatable">Not curatable</option>
  <option value="Uncertain">Uncertain</option>
  <option value="Not reviewed">Not reviewed</option>
</select>
:::
::: {.mb-3}
**Comment (optional)**  
<textarea id="comment" class="form-control" rows="3" placeholder="Evidence, edge case, missing field, false positive reason..." style="max-width: 520px;"></textarea>
:::
::: {.mb-3}
**BioAnalyzer version** (fixed for all submissions)  
<input type="text" id="bioanalyzer_version" class="form-control" readonly style="max-width: 200px; background-color: var(--bs-light, #f8f9fa);" title="Version is set at build time.">
:::

### Field-by-field validation (ground truth) {#field-validation-section}

For each field, provide the **curator TRUE label** (ground truth). Optionally mark whether BioAnalyzer's predicted status was correct.

::: {.grid}

```{r}
#| label: feedback-fields
#| echo: false
#| results: asis
for (col in STATUS_COLUMNS) {
  s <- safe_col(col)
  label <- gsub(" Status", "", col)
  cat(sprintf('
<div class="feedback-field">
  <strong>%s</strong>
  <p class="bioanalyzer-pred-line"><span id="pred_%s" class="bioanalyzer-pred text-muted">Select a PMID above and click Load row to see BioAnalyzer&rsquo;s prediction.</span></p>
  <small>Curator TRUE label for %s:</small>
  <select id="true_%s">
    <option value="Not reviewed">Not reviewed</option>
    <option value="ABSENT">ABSENT</option>
    <option value="PARTIALLY_PRESENT">PARTIALLY_PRESENT</option>
    <option value="PRESENT">PRESENT</option>
  </select>
  <small>Was BioAnalyzer correct for %s?</small>
  <select id="col_fb_%s">
    <option value="Not reviewed">Not reviewed</option>
    <option value="Correct">Correct</option>
    <option value="Incorrect">Incorrect</option>
    <option value="Unclear">Unclear</option>
  </select>
</div>
', label, s, label, s, label, s))
}
```

:::

**Submit review** — Opens GitHub to create an issue with your feedback (no file download). A bot will then open a Pull Request with the feedback CSV so the maintainer can view and download it (including your initials and comments).

::: {.mt-2}
<button type="button" id="btn-submit-review" class="btn btn-primary me-2">Submit review</button>
<button type="button" id="btn-download-csv" class="btn btn-outline-secondary">Download CSV only</button>
:::

::: {.mt-3 .pt-3 .border-top .small .text-muted}
**Notes:** Submissions are processed as issues and a tab will be opened for you to just click the create button. 
Reviewers will be looking into you submittd issue to evaluate the results
:::
:::
:::

::: {#thank-you-modal .thank-you-overlay}
::: {.thank-you-card}
**Thank you** for your time and input in helping to complete this review.  
Your feedback has been submitted; an issue and pull request are being created in the background.

Would you like to continue reviewing other papers?

::: {.thank-you-actions .mt-3}
<button type="button" id="btn-continue-review" class="btn btn-primary me-2">Yes</button>
<button type="button" id="btn-go-to-index" class="btn btn-outline-secondary">No</button>
:::
:::
:::

<script>
(function() {
  const schema = [
    "PMID","curator_id","overall_verdict","comment","timestamp","bioanalyzer_version",
    "pred__Host_Species_Status","pred__Body_Site_Status","pred__Condition_Status",
    "pred__Sequencing_Type_Status","pred__Taxa_Level_Status","pred__Sample_Size_Status",
    "true__Host_Species_Status","true__Body_Site_Status","true__Condition_Status",
    "true__Sequencing_Type_Status","true__Taxa_Level_Status","true__Sample_Size_Status",
    "col_feedback__Host_Species_Status","col_feedback__Body_Site_Status","col_feedback__Condition_Status",
    "col_feedback__Sequencing_Type_Status","col_feedback__Taxa_Level_Status","col_feedback__Sample_Size_Status"
  ];
  const statusCols = [
    "Host Species Status","Body Site Status","Condition Status",
    "Sequencing Type Status","Taxa Level Status","Sample Size Status"
  ];
  const safeCol = c => c.replace(/ /g, "_");

  let tableData = [];
  const el = document.getElementById("curator-table-data");
  if (el) try { tableData = JSON.parse(el.textContent); } catch (_) {}

  const configEl = document.getElementById("curator-config");
  let feedbackIssueUrl = "";
  if (configEl) {
    try {
      const cfg = JSON.parse(configEl.textContent);
      if (cfg.curator_id_default) document.getElementById("curator_id").value = cfg.curator_id_default;
      const verEl = document.getElementById("bioanalyzer_version");
      if (verEl) verEl.value = cfg.bioanalyzer_version_default || "1.0.0";
      if (cfg.feedback_issue_url) feedbackIssueUrl = cfg.feedback_issue_url;
    } catch (_) {}
  }
  if (!feedbackIssueUrl && typeof window !== "undefined") {
    if (window.location.hostname.endsWith(".github.io")) {
      var path = window.location.pathname.replace(/\/index\.html$/i, "").replace(/^\//, "").split("/").filter(Boolean);
      var repo = path[0];
      if (repo) {
        var user = window.location.hostname.replace(".github.io", "");
        feedbackIssueUrl = "https://github.com/" + user + "/" + repo + "/issues/new";
      }
    } else if (window.location.hostname === "waldronlab.io" && window.location.pathname.indexOf("curator-desk") !== -1) {
      feedbackIssueUrl = "https://github.com/waldronlab/curator-desk/issues/new";
    }
  }

  function setTitleFromPmid(pmid) {
    const row = tableData.find(r => String(r.PMID) === String(pmid));
    const titleEl = document.getElementById("title-prefill");
    if (row) {
      titleEl.textContent = row.Title || "(no title)";
      updatePredictionsForRow(row);
      const section = document.getElementById("field-validation-section");
      if (section) section.scrollIntoView({ behavior: "smooth", block: "start" });
    } else {
      titleEl.textContent = "(PMID not in table)";
      clearPredictions();
    }
  }

  function getPredictionValue(row, col) {
    if (!row) return "";
    const withSpaces = row[col];
    if (withSpaces !== undefined && withSpaces !== null && String(withSpaces).trim() !== "") return String(withSpaces).trim();
    const withUnderscores = row[col.replace(/\s+/g, "_")];
    if (withUnderscores !== undefined && withUnderscores !== null && String(withUnderscores).trim() !== "") return String(withUnderscores).trim();
    return "";
  }

  const predictionPlaceholder = "Select a PMID above and click Load row to see BioAnalyzer's prediction.";

  function updatePredictionsForRow(row) {
    statusCols.forEach(function(col) {
      const s = safeCol(col);
      const predEl = document.getElementById("pred_" + s);
      if (!predEl) return;
      const val = getPredictionValue(row, col);
      predEl.textContent = val ? "BioAnalyzer predicted: " + val : predictionPlaceholder;
      predEl.className = "bioanalyzer-pred " + (val ? "pred-" + val.toUpperCase().replace(/\s+/g, "_") : "text-muted");
    });
  }

  function clearPredictions() {
    statusCols.forEach(function(col) {
      const s = safeCol(col);
      const predEl = document.getElementById("pred_" + s);
      if (predEl) {
        predEl.textContent = predictionPlaceholder;
        predEl.className = "bioanalyzer-pred text-muted";
      }
    });
  }

  const quickSelect = document.getElementById("quick-select-pmid");
  if (quickSelect) {
    quickSelect.addEventListener("change", function() {
      const pmid = this.value;
      const fbPmid = document.getElementById("fb_pmid");
      if (fbPmid) fbPmid.value = pmid;
      if (pmid) setTitleFromPmid(pmid);
    });
  }

  const loadBtn = document.getElementById("btn-load-row");
  if (loadBtn) {
    loadBtn.addEventListener("click", function() {
      const pmid = document.getElementById("fb_pmid").value.trim();
      if (pmid) setTitleFromPmid(pmid);
    });
  }

  function escapeCsv(s) {
    if (s == null) return "";
    s = String(s);
    if (/[,"\n\r]/.test(s)) return '"' + s.replace(/"/g, '""') + '"';
    return s;
  }

  function buildFeedbackCsv() {
    const pmid = document.getElementById("fb_pmid").value.trim();
    const curatorId = document.getElementById("curator_id").value.trim();
    if (!pmid || !curatorId) {
      alert("Please fill Curator ID and PMID.");
      return null;
    }
    const row = tableData.find(r => String(r.PMID) === String(pmid));
    const pred = row || {};
    const ts = new Date().toISOString();
    const rowObj = {
      "PMID": pmid,
      "curator_id": curatorId,
      "overall_verdict": document.getElementById("overall_verdict").value,
      "comment": document.getElementById("comment").value.replace(/\r?\n/g, " "),
      "timestamp": ts,
      "bioanalyzer_version": document.getElementById("bioanalyzer_version").value
    };
    statusCols.forEach(col => {
      const s = safeCol(col);
      rowObj["pred__" + s] = (pred[col] !== undefined && pred[col] !== null) ? pred[col] : "";
      rowObj["true__" + s] = (document.getElementById("true_" + s) || {}).value || "Not reviewed";
      rowObj["col_feedback__" + s] = (document.getElementById("col_fb_" + s) || {}).value || "Not reviewed";
    });
    const header = schema.map(escapeCsv).join(",");
    const line = schema.map(k => escapeCsv(rowObj[k])).join(",");
    return { csv: header + "\n" + line + "\n", pmid: pmid };
  }

  function resetFormAfterSubmit() {
    const commentEl = document.getElementById("comment");
    if (commentEl) commentEl.value = "";
    statusCols.forEach(col => {
      const s = safeCol(col);
      const t = document.getElementById("true_" + s);
      const c = document.getElementById("col_fb_" + s);
      if (t) t.value = "Not reviewed";
      if (c) c.value = "Not reviewed";
    });
  }

  const downloadBtn = document.getElementById("btn-download-csv");
  if (downloadBtn) {
    downloadBtn.addEventListener("click", function() {
      const out = buildFeedbackCsv();
      if (!out) return;
      const blob = new Blob([out.csv], { type: "text/csv" });
      const a = document.createElement("a");
      a.href = URL.createObjectURL(blob);
      a.download = "curator_feedback_" + out.pmid + ".csv";
      a.click();
      URL.revokeObjectURL(a.href);
      resetFormAfterSubmit();
    });
  }

  const submitBtn = document.getElementById("btn-submit-review");
  if (submitBtn) {
    submitBtn.addEventListener("click", function() {
      const out = buildFeedbackCsv();
      if (!out) return;
      if (!feedbackIssueUrl) {
        alert("Submit review needs the GitHub repo URL. Set GITHUB_REPO when building (e.g. in your deploy workflow or locally: export GITHUB_REPO=https://github.com/owner/repo && quarto render).");
        return;
      }
      var bodyText = "Curator feedback submission. The CSV below contains the full feedback (curator initials, comments, field-by-field validation). Maintainers can review in this issue.\n\n```csv\n" + out.csv.trim() + "\n```";
      var title = "Curator feedback submission";
      var url = feedbackIssueUrl + "?title=" + encodeURIComponent(title) + "&body=" + encodeURIComponent(bodyText);
      window.open(url, "_blank");
      resetFormAfterSubmit();
      try { sessionStorage.setItem("curator-thank-you", "1"); } catch (e) {}
      window.location.reload();
    });
  }

  var thankYouModal = document.getElementById("thank-you-modal");
  if (thankYouModal) {
    try {
      if (sessionStorage.getItem("curator-thank-you") === "1") {
        sessionStorage.removeItem("curator-thank-you");
        thankYouModal.classList.add("show");
      }
    } catch (e) {}
    var btnYes = document.getElementById("btn-continue-review");
    var btnNo = document.getElementById("btn-go-to-index");
    if (btnYes) btnYes.addEventListener("click", function() {
      thankYouModal.classList.remove("show");
      var el = document.getElementById("candidate-papers");
      if (el) el.scrollIntoView({ behavior: "smooth", block: "start" });
    });
    if (btnNo) btnNo.addEventListener("click", function() {
      window.location.href = "./";
    });
  }
})();
</script>

<style>
.thank-you-overlay {
  position: fixed;
  top: 0; left: 0; right: 0; bottom: 0;
  background: rgba(0,0,0,0.5);
  display: flex;
  align-items: center;
  justify-content: center;
  z-index: 9999;
  opacity: 0;
  visibility: hidden;
  transition: opacity 0.25s ease, visibility 0.25s ease;
}
.thank-you-overlay.show {
  opacity: 1;
  visibility: visible;
}
.thank-you-card {
  background: var(--bs-body-bg, #fff);
  padding: 2rem;
  border-radius: 0.5rem;
  box-shadow: 0 0.5rem 2rem rgba(0,0,0,0.2);
  max-width: 420px;
  text-align: center;
}
.thank-you-actions { display: flex; justify-content: center; flex-wrap: wrap; gap: 0.5rem; }

#feedback-form .card-body { padding: 1.25rem 1.5rem; }
.feedback-field {
  margin-bottom: 1.25rem;
  padding: 0.75rem;
  background: var(--bs-light, #f8f9fa);
  border-radius: 0.35rem;
  border: 1px solid rgba(0,0,0,.06);
}
.feedback-field strong { display: block; margin-bottom: 0.35rem; }
.feedback-field .bioanalyzer-pred-line { margin: 0.35rem 0 0.5rem 0; min-height: 1.4em; }
.bioanalyzer-pred { font-size: 0.95em; font-weight: 500; }
.bioanalyzer-pred.pred-PRESENT { color: #198754; }
.bioanalyzer-pred.pred-ABSENT { color: #dc3545; }
.bioanalyzer-pred.pred-PARTIALLY_PRESENT { color: #fd7e14; }
.feedback-field select { margin-right: 0.5rem; margin-top: 0.25rem; max-width: 12em; }
.feedback-field small { color: var(--bs-secondary); }
@media (min-width: 768px) {
  .grid { display: grid; grid-template-columns: 1fr 1fr; gap: 1rem 1.5rem; }
}
</style>
