---
title: "Curator Desk"
format:
  html:
    toc: false
---

```{r echo = FALSE}
#| label: setup
#| include: false

# ---- Load dependencies ----
source("R/config.R")
source("R/data.R")
source("R/feedback.R")

# ---- Resolve data source ----

resolve_data_path <- function() {
  path <- Sys.getenv("CURATOR_DATA_URL", "")
  if (!nzchar(trimws(path))) path <- Sys.getenv("CURATOR_DATA_PATH", "")
  if (!nzchar(trimws(path))) path <- "data/sample.csv"
  path
}

data_path <- resolve_data_path()

# ---- Load and normalize dataset ----

raw_df <- load_data(data_path)
df <- normalize_dataset(raw_df)

# ---- Validate expected value columns ----

missing_values <- setdiff(VALUE_COLUMNS, names(df))

if (length(missing_values) > 0) {
  warning(
    sprintf(
      "Missing expected value columns: %s. The table and filtering may be partial.",
      paste(missing_values, collapse = ", ")
    )
  )
}

# ---- Hero stats ----

n_papers   <- nrow(df)
n_mapped   <- count_fully_mapped(df, ONTOLOGY_ID_COLUMNS)
n_bugsigdb <- if ("In bsgdb" %in% names(df)) sum(df[["In bsgdb"]], na.rm = TRUE) else 0L
n_new      <- n_papers - n_bugsigdb

repo_base <- sub("/issues/new$", "", CONFIG$feedback_issue_url)
```

```{r}
#| label: hero
#| echo: false
#| results: asis

github_btn <- if (nzchar(repo_base)) {
  sprintf('<a class="cd-btn cd-btn-ghost" href="%s" target="_blank" rel="noopener">View source on GitHub</a>', repo_base)
} else {
  ""
}

cat(sprintf('
<div class="cd-hero">
<div class="cd-hero-text">
<p class="cd-hero-kicker">BioAnalyzer &times; BugSigDB</p>
<h1 class="cd-hero-title">Curator Desk</h1>
<p class="cd-hero-sub">Review BioAnalyzer&rsquo;s extracted fields for candidate microbiome studies, confirm ontology mappings, record curator ground truth per PMID, and send validated rows straight into the BugSigDB curation pipeline.</p>
<div class="cd-hero-actions">
<a class="cd-btn cd-btn-primary" href="#candidate-papers">Browse candidate papers</a>
%s
</div>
</div>
<div class="cd-hero-stats">
<div class="cd-stat"><span class="cd-stat-value">%d</span><span class="cd-stat-label">Candidate papers</span></div>
<div class="cd-stat"><span class="cd-stat-value">%d</span><span class="cd-stat-label">Fully ontology-mapped</span></div>
<div class="cd-stat"><span class="cd-stat-value">%d</span><span class="cd-stat-label">Already in BugSigDB</span></div>
<div class="cd-stat"><span class="cd-stat-value">%d</span><span class="cd-stat-label">Awaiting curation</span></div>
</div>
</div>
', github_btn, n_papers, n_mapped, n_bugsigdb, n_new))
```

::: {.cd-row}

::: {.cd-col-main}

### How this works

::: {.cd-steps}
<div class="cd-step">
  <div class="cd-step-icon">1</div>
  <div><h4>Find papers quickly</h4><p>Search, sort, and filter the table below by any field, year, or BugSigDB status.</p></div>
</div>
<div class="cd-step">
  <div class="cd-step-icon">2</div>
  <div><h4>Record ground truth</h4><p>Open a PMID in the curator feedback form and enter the curator TRUE label, plus an optional &ldquo;was the prediction correct?&rdquo; flag, for each field.</p></div>
</div>
<div class="cd-step">
  <div class="cd-step-icon">3</div>
  <div><h4>Share feedback</h4><p>Submit straight to a GitHub issue with the feedback CSV pre-filled &mdash; no install, no login beyond GitHub.</p></div>
</div>
:::

:::

::: {.cd-col-side}

::: {.callout-note collapse="false"}
## Data Source & Setup

The table is built from data stored in the **`data/`** folder. To analyze your own papers:

1.  Place your CSV (or Parquet) file in `data/` (e.g., `data/sample.csv`).
2.  Run `quarto render`.

The system reads `data/sample.csv` by default. To use a different source, set the `CURATOR_DATA_PATH` environment variable to your specific path (e.g., `data/analyzed_papers.csv`) or use `CURATOR_DATA_URL` for remote files.

**Expected columns**

* **Required:** `PMID`
* **Recommended:** `Title`, `Journal`, `Year`
* **Value fields:** `Host Species`, `Body Site`, `Condition`, `Sample Size`, `Sequencing Type`
* **Ontology IDs (optional):** `Host Species Ontology ID`, `Body Site Ontology ID`, `Condition Ontology ID`
:::

:::

:::

::: {#candidate-papers .cd-card}

::: {.cd-card-header}
## Candidate Curatable Articles
<span class="cd-badge">`r n_papers` rows</span>
:::

```{r}
#| label: table-display
#| echo: false
#| warning: false
#| message: false
#| results: asis

library(dplyr)

cat('<div class="cd-table-wrap">')

if (nrow(df) == 0) {

  cat(
    "<div class='cd-card-body'>No data loaded. Set <code>CURATOR_DATA_URL</code> or <code>CURATOR_DATA_PATH</code>, ",
    "or add <code>data/sample.csv</code> and rebuild.</div>"
  )

} else {

  # ---- Select display columns ----
  # Ontology Candidates columns are deliberately excluded here - they're
  # picker-only metadata exported into the table-json blob below, not meant
  # to be read as a visible column.

  DISPLAY_COLUMNS <- c(
    "PubMed Link",
    "Title",
    "Journal",
    "Year",
    "Host Species",
    "Host Species Ontology ID",
    "Body Site",
    "Body Site Ontology ID",
    "Condition",
    "Condition Ontology ID",
    "Sample Size",
    "Sequencing Type",
    "Differential Abundance",
    "In bsgdb"
  )
  display_columns <- intersect(DISPLAY_COLUMNS, names(df))
  df_display <- df[, display_columns, drop = FALSE]

  # ---- Inform about missing value columns ----

  if (length(missing_values) > 0) {
    cat(sprintf(
      "<div class='cd-toolbar cd-toolbar-warn'>&#9888;&nbsp; Missing value columns: %s. The table may be partial.</div>",
      paste(missing_values, collapse = ", ")
    ))
  }

  cat('<div class="cd-toolbar"><strong>Tip:</strong>&nbsp; Use the search box and column filters to narrow your focus. An empty <strong>Ontology ID</strong> cell means that field needs a curator-confirmed mapping - see the feedback form below.</div>')

  # ---- Differential-abundance filter toggle ----
  # Per spec: curator-desk should default to showing only papers that report
  # differential abundance, with a "show all" escape hatch. Default to
  # unchecked when no row would match (e.g. the field is absent from this
  # dataset) so the table isn't silently emptied.
  diffab_index <- which(names(df_display) == "Differential Abundance")
  if (length(diffab_index) == 1) {
    n_diffab <- sum(df_display$`Differential Abundance`, na.rm = TRUE)
    checked_attr <- if (n_diffab > 0) "checked" else ""
    cat(sprintf('
<div class="cd-toolbar cd-filter-toolbar">
<label class="cd-filter-toggle">
<input type="checkbox" id="filter-diffab" %s>
Show only papers reporting differential abundance (%d of %d)
</label>
</div>
', checked_attr, n_diffab, nrow(df_display)))
  }

  # ---- Configure DataTable options ----

 dt_options <- list(
  pageLength = 50,
  lengthMenu = c(50, 100, 300, 500, 1000),
  scrollX = TRUE,
  stateSave = TRUE,
  fixedHeader = TRUE
)

  year_index <- which(names(df_display) == "Year")
  if (length(year_index) == 1) {
    dt_options$order <- list(list(year_index - 1, "desc"))
  }

  in_bsgdb_index <- which(names(df_display) == "In bsgdb")
  dt_column_defs <- list()
  if (length(in_bsgdb_index) == 1) {
    dt_column_defs <- append(
      dt_column_defs,
      list(
        list(
          targets = in_bsgdb_index - 1,
          title = "In bsgdb",
          render = DT::JS(
            "function(data, type, row) {",
            "  if (type === 'display') {",
            "    return data ? '<span class=\"bugsigdb-yes\">✓ Yes</span>' : '<span class=\"bugsigdb-no\">✗ No</span>';",
            "  }",
            "  return data;",
            "}"
          )
        )
      )
    )
  }

  dt_options$columnDefs <- dt_column_defs

  if (length(diffab_index) == 1) {
    dt_options$initComplete <- DT::JS(sprintf(
      "function(settings, json) {
        var api = this.api();
        var idx = %d;
        var checkbox = document.getElementById('filter-diffab');
        if (!checkbox) return;
        function apply() {
          api.column(idx).search(checkbox.checked ? '^true$' : '', true, false).draw();
        }
        checkbox.addEventListener('change', apply);
        apply();
      }",
      diffab_index - 1
    ))
  }

  # ---- Render table ----
  dt_table <- DT::datatable(
    df_display,
    filter = "top",
    extensions = c("FixedHeader"),
    options = dt_options,
    escape = FALSE,
    rownames = FALSE
  )

  if ("In bsgdb" %in% names(df_display)) {
    dt_table <- dt_table %>%
      DT::formatStyle(
        "In bsgdb",
        backgroundColor = DT::styleEqual(c(TRUE, FALSE), c("#e8f5e9", "#ffffff"))
      )
  }

  # Flag ontology-ID cells that are empty - a curator-confirmed mapping is
  # needed there (see the mapping picker in the feedback form below).
  for (onto_col in ONTOLOGY_ID_COLUMNS) {
    if (onto_col %in% names(df_display)) {
      dt_table <- dt_table %>%
        DT::formatStyle(
          onto_col,
          backgroundColor = DT::styleEqual("", "#fff9c4")
        )
    }
  }

  dt_table
}

cat('</div>')
```

:::

::: {#feedback-form-section .cd-card}

::: {.cd-card-header}
## Curator feedback
<span class="cd-badge cd-badge-muted">5 fields per paper</span>
:::

::: {.cd-card-body}

Use the table above to find papers (click a PMID's **PubMed Link** to open it). Then select or enter a PMID below; the page will scroll to the field-by-field validation form where you can enter your review. Feedback uses the same schema as the Streamlit app; because this site is static, **saving is done by downloading a CSV** per row&mdash;submit via your usual process (e.g. email or GitHub issue).

1. **Quick select** a PMID from the dropdown, or type a PMID and click **Load row** (the form will scroll into view).
2. For each field, review **BioAnalyzer predicted** and correct the **Curator value** if needed; for Host Species/Body Site/Condition, confirm or pick the **ontology mapping** from the suggested candidates (or enter one manually).
3. Click **Submit review** (or **Download CSV only**) to save one row.

```{r}
#| label: table-json
#| echo: false
#| results: asis
# Export table rows as JSON for the form's "Load row" button (PMID -> pred values)
if (nrow(df) > 0) {
  want <- c("PMID", "Title", VALUE_COLUMNS, ONTOLOGY_ID_COLUMNS, ONTOLOGY_CANDIDATES_COLUMNS)
  export_cols <- intersect(want, names(df))
  rows_json <- jsonlite::toJSON(
    df[, export_cols, drop = FALSE],
    auto_unbox = TRUE,
    dataframe = "rows"
  )
  cat(sprintf('<script type="application/json" id="curator-table-data">%s</script>', rows_json))
}
```

```{r}
#| label: curator-config
#| echo: false
#| results: asis
# Defaults for form (same as Python CONFIG)
cat(sprintf('<script type="application/json" id="curator-config">%s</script>',
            jsonlite::toJSON(list(curator_id_default = CONFIG$curator_id_default,
                                 bioanalyzer_version_default = CONFIG$bioanalyzer_version_default,
                                 feedback_issue_url = CONFIG$feedback_issue_url,
                                 # Single source of truth for the feedback form's fields: adding a
                                 # value/ontology-ID column to R/config.R is now enough,
                                 # js/feedback-form.js reads it from here instead of hardcoding
                                 # its own copy.
                                 value_columns = VALUE_COLUMNS,
                                 ontology_id_columns = ONTOLOGY_ID_COLUMNS,
                                 ontology_candidates_columns = ONTOLOGY_CANDIDATES_COLUMNS),
                            auto_unbox = TRUE)))
```

::: {#feedback-form .cd-form}

::: {.cd-form-row}
::: {.cd-form-field}
**Curator ID / initials**
<input type="text" id="curator_id" class="form-control" placeholder="e.g. Ronald Ouma">
:::
::: {.cd-form-field}
**BioAnalyzer version** (fixed for all submissions)
<input type="text" id="bioanalyzer_version" class="form-control" readonly title="Version is set at build time.">
:::
:::

::: {.cd-form-row}
::: {.cd-form-field .cd-form-field-wide}
**Quick select for feedback** — Select a PMID from the table, or type below.

```{r}
#| label: quick-select-pmid
#| echo: false
#| results: asis
if (nrow(df) > 0 && "PMID" %in% names(df)) {
  opts <- df[, c("PMID", if ("Title" %in% names(df)) "Title" else NULL), drop = FALSE]
  cat('<select id="quick-select-pmid" class="form-select"><option value="">— Select a PMID —</option>')
  for (i in seq_len(nrow(opts))) {
    pm <- opts$PMID[i]
    lbl <- if ("Title" %in% names(opts)) paste0(pm, " — ", substr(opts$Title[i], 1, 50), if (nchar(as.character(opts$Title[i])) > 50) "…" else "") else as.character(pm)
    lbl <- gsub("&", "&amp;", gsub("<", "&lt;", gsub(">", "&gt;", lbl)))
    cat(sprintf('<option value="%s">%s</option>', pm, lbl))
  }
  cat('</select>')
}
```
:::
::: {.cd-form-field}
**PMID**
<div class="input-group">
  <input type="text" id="fb_pmid" class="form-control" placeholder="e.g. 31215600">
  <button type="button" id="btn-load-row" class="btn btn-outline-secondary">Load row</button>
</div>
:::
:::

::: {.cd-form-row}
::: {.cd-form-field .cd-form-field-wide}
**Title (from table)**
<div class="cd-title-prefill"><span id="title-prefill" class="text-muted fst-italic">Select a PMID above and click Load row.</span></div>
:::
::: {.cd-form-field}
**Overall paper verdict**
<select id="overall_verdict" class="form-select">
  <option value="Curatable">Curatable</option>
  <option value="Not curatable">Not curatable</option>
  <option value="Uncertain">Uncertain</option>
  <option value="Not reviewed">Not reviewed</option>
</select>
:::
:::

::: {.cd-form-row}
::: {.cd-form-field .cd-form-field-wide}
**Comment (optional)**
<textarea id="comment" class="form-control" rows="2" placeholder="Evidence, edge case, missing field, false positive reason..."></textarea>
:::
:::

### Field-by-field validation (ground truth) {#field-validation-section}

For each field, review **BioAnalyzer's predicted value** and correct it if needed. For Host Species, Body Site, and Condition, also confirm or pick the **ontology mapping** &mdash; a dropdown of BioAnalyzer's suggested candidates appears when a field's mapping wasn't auto-applied (see the **Ontology ID** column in the table above); otherwise enter one manually.

::: {.grid}

```{r}
#| label: feedback-fields
#| echo: false
#| results: asis
onto_col_for <- function(value_col) {
  candidate <- paste(value_col, "Ontology ID")
  if (candidate %in% ONTOLOGY_ID_COLUMNS) candidate else NA_character_
}

for (col in VALUE_COLUMNS) {
  s <- safe_col(col)
  onto_col <- onto_col_for(col)
  onto_block <- ""
  if (!is.na(onto_col)) {
    onto_s <- safe_col(onto_col)
    onto_block <- sprintf('
<p class="bioanalyzer-pred-line"><span id="pred_%s" class="bioanalyzer-pred text-muted">Ontology ID: (load a row to see it)</span></p>
<small>Ontology mapping for %s:</small>
<select id="true_%s" class="cd-onto-select">
<option value="">Not reviewed</option>
</select>
<input type="text" id="true_%s_manual" class="form-control cd-onto-manual" placeholder="Enter ontology ID manually (e.g. NCBITaxon:9606)" style="display:none;">
', onto_s, col, onto_s, onto_s)
  }
  cat(sprintf('
<div class="feedback-field">
<strong>%s</strong>
<p class="bioanalyzer-pred-line"><span id="pred_%s" class="bioanalyzer-pred text-muted">Select a PMID above and click Load row to see BioAnalyzer&rsquo;s prediction.</span></p>
<small>Curator value for %s:</small>
<input type="text" id="true_%s" class="form-control">
<small>Was BioAnalyzer correct for %s?</small>
<select id="col_fb_%s">
<option value="Not reviewed">Not reviewed</option>
<option value="Correct">Correct</option>
<option value="Incorrect">Incorrect</option>
<option value="Unclear">Unclear</option>
</select>
%s</div>
', col, s, col, s, col, s, onto_block))
}
```

:::

::: {.cd-submit-row}
<div class="cd-submit-note"><strong>Submit review</strong> opens GitHub to create an issue with your feedback CSV in the body. A bot acknowledges the issue automatically; maintainers review the CSV directly in the issue.</div>
<div class="cd-submit-actions">
  <button type="button" id="btn-submit-review" class="btn btn-primary">Submit review</button>
  <button type="button" id="btn-download-csv" class="btn btn-outline-secondary">Download CSV only</button>
</div>
:::

::: {.cd-form-footnote}
**Notes:** Submissions are processed as issues and a tab will be opened for you to just click the create button. Reviewers will look into your submitted issue to evaluate the results.
:::

:::

:::

:::

```{r}
#| label: footer
#| echo: false
#| results: asis

repo_link  <- if (nzchar(repo_base)) sprintf('<a href="%s" target="_blank" rel="noopener">GitHub repository</a>', repo_base) else ""
issue_link <- if (nzchar(CONFIG$feedback_issue_url)) sprintf('<a href="%s" target="_blank" rel="noopener">Open an issue</a>', CONFIG$feedback_issue_url) else ""

links <- paste(
  Filter(nzchar, c(
    repo_link,
    issue_link,
    '<a href="https://bugsigdb.org" target="_blank" rel="noopener">BugSigDB</a>',
    '<a href="https://pubmed.ncbi.nlm.nih.gov" target="_blank" rel="noopener">PubMed</a>'
  )),
  collapse = ""
)

cat(sprintf('
<div class="cd-footer">
<div>Curator Desk &middot; BioAnalyzer v%s &middot; PubMed &rarr; BioAnalyzer &rarr; Curator Desk &rarr; BugSigDB</div>
<div class="cd-footer-links">%s</div>
</div>
', CONFIG$bioanalyzer_version_default, links))
```

::: {#thank-you-modal .thank-you-overlay}
::: {.thank-you-card}
**Thank you** for your time and input in helping to complete this review.
Your feedback has been submitted; an issue and pull request are being created in the background.

Would you like to continue reviewing other papers?

::: {.thank-you-actions .mt-3}
<button type="button" id="btn-continue-review" class="btn btn-primary me-2">Yes</button>
<button type="button" id="btn-go-to-index" class="btn btn-outline-secondary">No</button>
:::
:::
:::

<script src="js/feedback-form.js"></script>
