---
title: "Curator Desk"
format:
  html:
    toc: false
---

```{r echo = FALSE}
#| label: setup
#| include: false

# ---- Load dependencies ----
source("R/config.R")
source("R/data.R")
source("R/feedback.R")

# ---- Resolve data source ----

resolve_data_path <- function() {
  path <- Sys.getenv("CURATOR_DATA_URL", "")
  if (!nzchar(trimws(path))) path <- Sys.getenv("CURATOR_DATA_PATH", "")
  if (!nzchar(trimws(path))) path <- "data/sample.csv"
  path
}

data_path <- resolve_data_path()

# ---- Load and normalize dataset ----

raw_df <- load_data(data_path)
df <- normalize_dataset(raw_df)

# ---- Validate expected status columns ----

missing_status <- setdiff(STATUS_COLUMNS, names(df))

if (length(missing_status) > 0) {
  warning(
    sprintf(
      "Missing expected status columns: %s. Priority and filtering may be partial.",
      paste(missing_status, collapse = ", ")
    )
  )
}

# ---- Hero stats ----

n_papers   <- nrow(df)
n_high     <- if ("Priority" %in% names(df)) sum(df$Priority >= 4, na.rm = TRUE) else 0L
n_bugsigdb <- if ("in_bugsigdb" %in% names(df)) sum(df$in_bugsigdb, na.rm = TRUE) else 0L
n_new      <- n_papers - n_bugsigdb

repo_base <- sub("/issues/new$", "", CONFIG$feedback_issue_url)
```

```{r}
#| label: hero
#| echo: false
#| results: asis

github_btn <- if (nzchar(repo_base)) {
  sprintf('<a class="cd-btn cd-btn-ghost" href="%s" target="_blank" rel="noopener">View source on GitHub</a>', repo_base)
} else {
  ""
}

cat(sprintf('
<div class="cd-hero">
<div class="cd-hero-text">
<p class="cd-hero-kicker">BioAnalyzer &times; BugSigDB</p>
<h1 class="cd-hero-title">Curator Desk</h1>
<p class="cd-hero-sub">Review BioAnalyzer&rsquo;s predicted statuses for candidate microbiome studies, record curator ground truth per PMID, and send validated rows straight into the BugSigDB curation pipeline.</p>
<div class="cd-hero-actions">
<a class="cd-btn cd-btn-primary" href="#candidate-papers">Browse candidate papers</a>
%s
</div>
</div>
<div class="cd-hero-stats">
<div class="cd-stat"><span class="cd-stat-value">%d</span><span class="cd-stat-label">Candidate papers</span></div>
<div class="cd-stat"><span class="cd-stat-value">%d</span><span class="cd-stat-label">High priority</span></div>
<div class="cd-stat"><span class="cd-stat-value">%d</span><span class="cd-stat-label">Already in BugSigDB</span></div>
<div class="cd-stat"><span class="cd-stat-value">%d</span><span class="cd-stat-label">Awaiting curation</span></div>
</div>
</div>
', github_btn, n_papers, n_high, n_bugsigdb, n_new))
```

::: {.cd-row}

::: {.cd-col-main}

### How this works

::: {.cd-steps}
<div class="cd-step">
  <div class="cd-step-icon">1</div>
  <div><h4>Find papers quickly</h4><p>Search, sort, and filter the table below &mdash; the <strong>Priority</strong> score surfaces the highest-value candidates first.</p></div>
</div>
<div class="cd-step">
  <div class="cd-step-icon">2</div>
  <div><h4>Record ground truth</h4><p>Open a PMID in the curator feedback form and enter the curator TRUE label, plus an optional &ldquo;was the prediction correct?&rdquo; flag, for each field.</p></div>
</div>
<div class="cd-step">
  <div class="cd-step-icon">3</div>
  <div><h4>Share feedback</h4><p>Submit straight to a GitHub issue with the feedback CSV pre-filled &mdash; no install, no login beyond GitHub.</p></div>
</div>
:::

:::

::: {.cd-col-side}

::: {.callout-note collapse="false"}
## Data Source & Setup

The table is built from data stored in the **`data/`** folder. To analyze your own papers:

1.  Place your CSV (or Parquet) file in `data/` (e.g., `data/sample.csv`).
2.  Run `quarto render`.

The system reads `data/sample.csv` by default. To use a different source, set the `CURATOR_DATA_PATH` environment variable to your specific path (e.g., `data/analyzed_papers.csv`) or use `CURATOR_DATA_URL` for remote files.

**Expected columns**

* **Required:** `PMID`
* **Recommended:** `Title`, `Journal`, `Year`, `Summary`
* **Status values:** `ABSENT` | `PARTIALLY_PRESENT` | `PRESENT`
:::

:::

:::

::: {#candidate-papers .cd-card}

::: {.cd-card-header}
## Candidate Curatable Articles
<span class="cd-badge">`r n_papers` rows</span>
:::

```{r}
#| label: table-display
#| echo: false
#| warning: false
#| message: false
#| results: asis

library(dplyr)

cat('<div class="cd-table-wrap">')

if (nrow(df) == 0) {

  cat(
    "<div class='cd-card-body'>No data loaded. Set <code>CURATOR_DATA_URL</code> or <code>CURATOR_DATA_PATH</code>, ",
    "or add <code>data/sample.csv</code> and rebuild.</div>"
  )

} else {

  # ---- Inform about missing status columns ----

  if (length(missing_status) > 0) {
    cat(sprintf(
      "<div class='cd-toolbar cd-toolbar-warn'>&#9888;&nbsp; Missing status columns: %s. Priority and filtering may be partial.</div>",
      paste(missing_status, collapse = ", ")
    ))
  }

  cat('<div class="cd-toolbar"><strong>Tip:</strong>&nbsp; Sort by <strong>Priority</strong> (descending) to review the most promising candidates first. Use the search box and column filters to narrow your focus.</div>')

  # ---- Select display columns ----

  DISPLAY_COLUMNS <- c(
    "PubMed Link",
    "Title",
    "Journal",
    "Year",
    "Host Species",
    "Host Species Status",
    "Body Site",
    "Body Site Status",
    "Condition",
    "Condition Status",
    "Sequencing Type",
    "Sequencing Type Status",
    "Sample Size",
    "Sample Size Status",
    "has_differential_abundance",
    "differential_abundance_confidence",
    "in_bugsigdb",
    "Priority"
  )
  display_columns <- intersect(DISPLAY_COLUMNS, names(df))
  df_display <- df[, display_columns, drop = FALSE]

  # ---- Configure DataTable options ----

 dt_options <- list(
  pageLength = 50,
  lengthMenu = c(50, 100, 300, 500, 1000),
  scrollX = TRUE,
  stateSave = TRUE,
  fixedHeader = TRUE
)

  priority_index <- which(names(df_display) == "Priority")

  if (length(priority_index) == 1) {
    dt_options$order <- list(list(priority_index - 1, "desc"))
  }

  in_bugsigdb_index <- which(names(df_display) == "in_bugsigdb")
  dt_column_defs <- list()
  if (length(in_bugsigdb_index) == 1) {
    dt_column_defs <- append(
      dt_column_defs,
      list(
        list(
          targets = in_bugsigdb_index - 1,
          title = "In BugSigDB",
          render = DT::JS(
            "function(data, type, row) {",
            "  if (type === 'display') {",
            "    return data ? '<span class=\"bugsigdb-yes\">✓ Yes</span>' : '<span class=\"bugsigdb-no\">✗ No</span>';",
            "  }",
            "  return data;",
            "}"
          )
        )
      )
    )
  }

  dt_options$columnDefs <- dt_column_defs

  # ---- Render table ----
  dt_table <- DT::datatable(
    df_display,
    filter = "top",
    extensions = c("FixedHeader"),
    options = dt_options,
    escape = FALSE,
    rownames = FALSE
  )

  if ("in_bugsigdb" %in% names(df_display)) {
    dt_table <- dt_table %>%
      DT::formatStyle(
        "in_bugsigdb",
        backgroundColor = DT::styleEqual(c(TRUE, FALSE), c("#e8f5e9", "#ffffff"))
      )
  }

  for (status_col in STATUS_COLUMNS) {
    if (status_col %in% names(df_display)) {
      dt_table <- dt_table %>%
        DT::formatStyle(
          status_col,
          backgroundColor = DT::styleEqual(
            c("PRESENT", "PARTIALLY_PRESENT", "ABSENT"),
            c("#c8e6c9", "#fff9c4", "#ffcdd2")
          )
        )
    }
  }

  dt_table
}

cat('</div>')
```

:::

::: {#feedback-form-section .cd-card}

::: {.cd-card-header}
## Curator feedback
<span class="cd-badge cd-badge-muted">5 fields per paper</span>
:::

::: {.cd-card-body}

Use the table above to find papers (the **Open** button in the **Published** column opens the paper on PubMed). Then select or enter a PMID below; the page will scroll to the field-by-field validation form where you can enter your review. Feedback uses the same schema as the Streamlit app; because this site is static, **saving is done by downloading a CSV** per row&mdash;submit via your usual process (e.g. email or GitHub issue).

1. **Quick select** a PMID from the dropdown, or type a PMID and click **Load row** (the form will scroll into view).
2. For each field, review **BioAnalyzer predicted** and set **Curator TRUE label** and **Was BioAnalyzer correct?**
3. Click **Submit review** (or **Download CSV only**) to save one row.

```{r}
#| label: table-json
#| echo: false
#| results: asis
# Export table rows as JSON for the form's "Load row" button (PMID -> pred values)
if (nrow(df) > 0) {
  want <- c("PMID", "Title", STATUS_COLUMNS)
  export_cols <- intersect(want, names(df))
  rows_json <- jsonlite::toJSON(
    df[, export_cols, drop = FALSE],
    auto_unbox = TRUE,
    dataframe = "rows"
  )
  cat(sprintf('<script type="application/json" id="curator-table-data">%s</script>', rows_json))
}
```

```{r}
#| label: curator-config
#| echo: false
#| results: asis
# Defaults for form (same as Python CONFIG)
cat(sprintf('<script type="application/json" id="curator-config">%s</script>',
            jsonlite::toJSON(list(curator_id_default = CONFIG$curator_id_default,
                                 bioanalyzer_version_default = CONFIG$bioanalyzer_version_default,
                                 feedback_issue_url = CONFIG$feedback_issue_url),
                            auto_unbox = TRUE)))
```

::: {#feedback-form .cd-form}

::: {.cd-form-row}
::: {.cd-form-field}
**Curator ID / initials**
<input type="text" id="curator_id" class="form-control" placeholder="e.g. Ronald Ouma">
:::
::: {.cd-form-field}
**BioAnalyzer version** (fixed for all submissions)
<input type="text" id="bioanalyzer_version" class="form-control" readonly title="Version is set at build time.">
:::
:::

::: {.cd-form-row}
::: {.cd-form-field .cd-form-field-wide}
**Quick select for feedback** — Select a PMID from the table, or type below.

```{r}
#| label: quick-select-pmid
#| echo: false
#| results: asis
if (nrow(df) > 0 && "PMID" %in% names(df)) {
  opts <- df[, c("PMID", if ("Title" %in% names(df)) "Title" else NULL), drop = FALSE]
  cat('<select id="quick-select-pmid" class="form-select"><option value="">— Select a PMID —</option>')
  for (i in seq_len(nrow(opts))) {
    pm <- opts$PMID[i]
    lbl <- if ("Title" %in% names(opts)) paste0(pm, " — ", substr(opts$Title[i], 1, 50), if (nchar(as.character(opts$Title[i])) > 50) "…" else "") else as.character(pm)
    lbl <- gsub("&", "&amp;", gsub("<", "&lt;", gsub(">", "&gt;", lbl)))
    cat(sprintf('<option value="%s">%s</option>', pm, lbl))
  }
  cat('</select>')
}
```
:::
::: {.cd-form-field}
**PMID**
<div class="input-group">
  <input type="text" id="fb_pmid" class="form-control" placeholder="e.g. 31215600">
  <button type="button" id="btn-load-row" class="btn btn-outline-secondary">Load row</button>
</div>
:::
:::

::: {.cd-form-row}
::: {.cd-form-field .cd-form-field-wide}
**Title (from table)**
<div class="cd-title-prefill"><span id="title-prefill" class="text-muted fst-italic">Select a PMID above and click Load row.</span></div>
:::
::: {.cd-form-field}
**Overall paper verdict**
<select id="overall_verdict" class="form-select">
  <option value="Curatable">Curatable</option>
  <option value="Not curatable">Not curatable</option>
  <option value="Uncertain">Uncertain</option>
  <option value="Not reviewed">Not reviewed</option>
</select>
:::
:::

::: {.cd-form-row}
::: {.cd-form-field .cd-form-field-wide}
**Comment (optional)**
<textarea id="comment" class="form-control" rows="2" placeholder="Evidence, edge case, missing field, false positive reason..."></textarea>
:::
:::

### Field-by-field validation (ground truth) {#field-validation-section}

For each field, provide the **curator TRUE label** (ground truth). Optionally mark whether BioAnalyzer's predicted status was correct.

::: {.grid}

```{r}
#| label: feedback-fields
#| echo: false
#| results: asis
for (col in STATUS_COLUMNS) {
  s <- safe_col(col)
  label <- gsub(" Status", "", col)
  cat(sprintf('
<div class="feedback-field">
<strong>%s</strong>
<p class="bioanalyzer-pred-line"><span id="pred_%s" class="bioanalyzer-pred text-muted">Select a PMID above and click Load row to see BioAnalyzer&rsquo;s prediction.</span></p>
<small>Curator TRUE label for %s:</small>
<select id="true_%s">
<option value="Not reviewed">Not reviewed</option>
<option value="ABSENT">ABSENT</option>
<option value="PARTIALLY_PRESENT">PARTIALLY_PRESENT</option>
<option value="PRESENT">PRESENT</option>
</select>
<small>Was BioAnalyzer correct for %s?</small>
<select id="col_fb_%s">
<option value="Not reviewed">Not reviewed</option>
<option value="Correct">Correct</option>
<option value="Incorrect">Incorrect</option>
<option value="Unclear">Unclear</option>
</select>
</div>
', label, s, label, s, label, s))
}
```

:::

::: {.cd-submit-row}
<div class="cd-submit-note"><strong>Submit review</strong> opens GitHub to create an issue with your feedback. A bot then opens a Pull Request with the feedback CSV so the maintainer can view and download it.</div>
<div class="cd-submit-actions">
  <button type="button" id="btn-submit-review" class="btn btn-primary">Submit review</button>
  <button type="button" id="btn-download-csv" class="btn btn-outline-secondary">Download CSV only</button>
</div>
:::

::: {.cd-form-footnote}
**Notes:** Submissions are processed as issues and a tab will be opened for you to just click the create button. Reviewers will look into your submitted issue to evaluate the results.
:::

:::

:::

:::

```{r}
#| label: footer
#| echo: false
#| results: asis

repo_link  <- if (nzchar(repo_base)) sprintf('<a href="%s" target="_blank" rel="noopener">GitHub repository</a>', repo_base) else ""
issue_link <- if (nzchar(CONFIG$feedback_issue_url)) sprintf('<a href="%s" target="_blank" rel="noopener">Open an issue</a>', CONFIG$feedback_issue_url) else ""

links <- paste(
  Filter(nzchar, c(
    repo_link,
    issue_link,
    '<a href="https://bugsigdb.org" target="_blank" rel="noopener">BugSigDB</a>',
    '<a href="https://pubmed.ncbi.nlm.nih.gov" target="_blank" rel="noopener">PubMed</a>'
  )),
  collapse = ""
)

cat(sprintf('
<div class="cd-footer">
<div>Curator Desk &middot; BioAnalyzer v%s &middot; PubMed &rarr; BioAnalyzer &rarr; Curator Desk &rarr; BugSigDB</div>
<div class="cd-footer-links">%s</div>
</div>
', CONFIG$bioanalyzer_version_default, links))
```

::: {#thank-you-modal .thank-you-overlay}
::: {.thank-you-card}
**Thank you** for your time and input in helping to complete this review.
Your feedback has been submitted; an issue and pull request are being created in the background.

Would you like to continue reviewing other papers?

::: {.thank-you-actions .mt-3}
<button type="button" id="btn-continue-review" class="btn btn-primary me-2">Yes</button>
<button type="button" id="btn-go-to-index" class="btn btn-outline-secondary">No</button>
:::
:::
:::

<script src="js/feedback-form.js"></script>
